---
title: "Predicting Troops Data Assembly"
author: "Michael E Flynn"
date: "3/17/2021"
output: html_document
---


```{r setup, echo = FALSE, include = FALSE, warning = FALSE}
library(tidyverse)
library(troopdata)
library(peacesciencer)
library(cshapes)
library(ggtext)
library(broom)
library(countrycode)
library(here)
library(tictoc)
library(tidyquant)
library(purrr)

#devtools::install_github("svmiller/peacesciencer")

knitr::opts_chunk$set(echo = FALSE, dpi = 400, fig.height = 5, fig.width = 8)

theme_flynn <- theme_linedraw() + theme(text = element_markdown(size = 11),
                                        plot.title = element_markdown(size = 16),
                                        plot.subtitle = element_markdown(size = 12),
                                        plot.caption = element_markdown(face = "italic", size = 8),
                                        strip.background = element_rect(fill = "gray80", color = "black"),
                                        strip.text = element_text(color = "black", face = "bold"),
                                        panel.grid.major = element_line(color = "gray70", size = 0.5),
                                        panel.grid.minor = element_line(color = "gray90", size = 0.25),
                                        axis.title = element_text(face = "bold", size = 12),
                                        axis.title.y = element_text(margin = margin(t = 0, r = 0.5, b = 0, l = 0, unit = "cm")),
                                        axis.title.x = element_text(margin = margin(t = 0.5, r = 0, b = 0, l = 0, unit = "cm")),
                                        legend.title = element_text(face = "bold"))

```


# Read in data for assembly 

## Included data

1. Uses Miller's `{peacesciencer}` to generate the following data:
  - ATOP alliances
  - V-Dem data for democracy
  - COW NMC data
  - Gleditsch minimum distance data between dyads
  - COW Trade data v4.0
  - GDP data from Anders, Therese, Christopher J. Fariss, and Jonathan N. Markowitz. 2020. "Bread Before Guns or Butter: Introducing Surplus Domestic Product (SDP)" International Studies Quarterly 64(2): 392–405.
2. CPI data from FRED. Original data from US Bureau of Labor Statistics
3. US conflict data is from PRIO armed conflict dataset
4. Uses `cshapes` data from Wiedmann and Gleditsch to create minimum distance variables and spatially weighted variables.
5. SIPRI defense burden data from 1949-2019
6. Voeten's UN ideal point data to calculate ideal point distance values
7. Troop deployment data from `{troopdata}` package.
8. Major power troop deployment data from Braithewaite.
9. IMF DOTS for trade data from 2014-Present.


## Notes:

1. We treat West Germany and the United Germany as a single polity. For 1990 where there is an overlap in values we take the highest value for any variable. 
2. Note that the GDP and population data need to be handled separately. The authors use the Gleditsch and Ward country codes. If you use the COW codes to generate the base data then some countries will be missing. For example, about half of Germany's observations are missing if you use the COW system.

```{r data assembly, echo=FALSE, message=FALSE}

#### CPI Data for Inflation adjustments ####
# Set to 2015 prices as baseline
cpi <- tidyquant::tq_get("CUUR0000SA0R", "economic.data", from = "1950-01-01") %>%
  mutate(year = format(as.Date(date, "%Y-%m-%d"), "%Y")) %>%
  group_by(year) %>%
  dplyr::summarise(cpi = mean(price)) %>%
  mutate(cpi = cpi/cpi[year==2015],
         year = as.numeric(year))

dyads <- peacesciencer::create_dyadyears(system = "cow", directed = TRUE) %>% 
  add_atop_alliance() %>% 
  add_democracy() %>% 
  add_minimum_distance(system = "cow") %>% 
  add_nmc() %>% 
  filter(ccode1 == 2 | ccode1 == 365) %>% 
  filter(year >= 1950) %>% 
  mutate(ccode2 = ifelse(ccode2 == 260, 255, ccode2)) 


germany.1990 <- dyads %>% 
  filter(ccode2 == 255 & year == 1990) %>% 
  group_by(ccode1, ccode2, year) %>% 
  dplyr::summarise(across(everything(), ~max(., na.rm = TRUE)))

dyads.complete <- dyads %>% 
  dplyr::filter(!(ccode2 == 255 & year == 1990)) %>% 
  bind_rows(germany.1990) %>% 
  arrange(ccode1, ccode2, year)


# Have to create the GDP data separately because they use different codes for Germany. Using COW system creates missing values.
# Using GW codes includes the appropriate values.
gdp <- peacesciencer::create_stateyears(system = "gw") %>% 
  add_sdp_gdp(system = "gw") %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode1" = "gwcode") %>% 
  mutate(ccode1 = ifelse(ccode1 == 260, 255, ccode1))

gdp.sp <- peacesciencer::create_stateyears(system = "gw") %>% 
  add_sdp_gdp(system = "gw") %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode2" = "gwcode") %>% 
  mutate(ccode2 = ifelse(ccode2 == 260, 255, ccode2)) 



# Separate democracy index for spatial measures
dem.sp <- peacesciencer::create_stateyears(system = "cow") %>% 
  add_democracy() %>% 
  filter(year >= 1950) %>% 
  group_by(ccode, year) %>%
  mutate(ccode = ifelse(ccode == 260, 255, ccode)) %>% 
  dplyr::summarise(across(c("v2x_polyarchy", "polity2"), ~max(.x, na.rm = TRUE))) %>% 
  dplyr::rename("ccode2" = "ccode")
  


#### Gleditsch Trade Data ####
trade.gled <- read_delim(here::here("../Data/trade_dd.asc"), delim = " ") %>% 
  filter(numa %in% c(2, 365)) %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode1" = "numa", "ccode2" = "numb", "imports_to_ccode1_g" = "impab",  "exports_from_ccode1_g" = "expab") %>% 
  mutate(ccode2 = ifelse(ccode2 == 260, 255, ccode2)) 
  


#### COW Trade Data ####
trade.cow.data <- read_csv(here("../Data", "Dyadic_COW_4.0.csv")) %>%
  filter(year >= 1950 & ccode1 %in% c(2, 365)) %>%
  dplyr::mutate(flow1 = ifelse(flow1 == -9, NA, flow1),
                flow2 = ifelse(flow2 == -9, NA, flow2)) %>% 
  dplyr::rename("imports_to_ccode1_cow" = "flow1",
                "exports_from_ccode1_cow" = "flow2") %>%
  select("ccode1", "ccode2","year", "imports_to_ccode1_cow", "exports_from_ccode1_cow") %>% 
  mutate(ccode2 = case_when(
    ccode2 == 260 ~ 255,
    TRUE ~ ccode2
  )) 


# Combine trade data for COW and Gleditsch
trade.com <- trade.cow.data %>% 
  full_join(trade.gled) %>% 
  mutate(exports_from_ccode1 = case_when(
    is.na(exports_from_ccode1_cow) ~ exports_from_ccode1_g,
    TRUE ~ (exports_from_ccode1_cow)
  ),
  imports_to_ccode1 = case_when(
    is.na(imports_to_ccode1_cow) ~ imports_to_ccode1_g,
    TRUE ~ (imports_to_ccode1_cow)
  ),
  across(starts_with("imports|exports")), ~round(digits = 4))



#### US IMF Trade Data ####
# WARNING!!! for whatever reason the pivot function sometimes doesn't work on every other pass.
# Make sure this is in long form

# Also, LOTS of stupid names match the Netherlands country codes. Filter those out.
netherlandsfilter <- c("Curaçao, Kingdom of the Netherlands", "Sint Maarten, Kingdom of the Netherlands", "Aruba, Kingdom of the Netherlands")

`%notin%` <- Negate(`%in%`) # create this to get rid of lots of duplicate Netherlands values. Appear as a result of protectorates or territories.

# IMF DOTS Exports to counterpart countries from US
exports.us <- readxl::read_xlsx(here::here("../Data", "Exports_to_Counterpart_Countries.xlsx"), skip = 5)

exports.us <- exports.us %>%
  dplyr::rename("country" = 1) %>%
  pivot_longer(cols = 2:ncol(exports.us),
               values_to = "exports",
               names_to = "year") %>%
  filter(country %notin% netherlandsfilter) %>%
  mutate(ccode = countrycode(country, "country.name", "cown"),
         ccode = ifelse(country == "Eastern Germany", 265, ccode),
         ccode = ifelse(country == "China, P.R.", 710, ccode),
         ccode = ifelse(country == "Central African Rep.", 482, ccode),
         year = as.numeric(year)) %>%
  filter(year > 2014)  %>%
  mutate(ccode1 = 2) %>% 
  dplyr::rename("exports_from_ccode1" = "exports", "ccode2" = "ccode") %>%
  dplyr::select("ccode1", "ccode2", "year", "exports_from_ccode1")



# IMF DOTS Imports from counterpart countries into US
imports.us <- readxl::read_xlsx(here("../Data", "Imports_from_Counterpart_Countries.xlsx"), skip = 5)

imports.us <- imports.us %>%
  dplyr::rename("country" = 1) %>%
  pivot_longer(cols = 2:ncol(imports.us),
               values_to = "imports",
               names_to = "year") %>%
  filter(country %notin% netherlandsfilter) %>%
  mutate(ccode = countrycode(country, "country.name", "cown"),
         ccode = ifelse(country == "Eastern Germany", 265, ccode),
         ccode = ifelse(country == "China, P.R.", 710, ccode),
         ccode = ifelse(country == "Central African Rep.", 482, ccode),
         year = as.numeric(year)) %>%
  filter(year > 2014) %>%
  mutate(ccode1 = 2) %>% 
  dplyr::rename("imports_to_ccode1" = "imports", "ccode2" = "ccode") %>%
  dplyr::select("ccode1", "ccode2", "year", "imports_to_ccode1")


# Combine US data
imf.us <- exports.us %>%
  full_join(imports.us, by = c("ccode1", "ccode2", "year"))




#### Russian Trade Data ####
exports.russia <- readxl::read_xlsx(here::here("../Data", "Exports_and_Imports_by_areas_and_Co (2).xlsx"), sheet = 1, skip = 5)

exports.russia <- exports.russia %>%
  dplyr::rename("country" = 1) %>%
  pivot_longer(cols = 2:ncol(exports.russia),
               values_to = "exports",
               names_to = "year") %>%
  filter(country %notin% netherlandsfilter) %>%
  mutate(ccode = countrycode(country, "country.name", "cown"),
         ccode = ifelse(country == "Eastern Germany", 265, ccode),
         ccode = ifelse(country == "China, P.R.", 710, ccode),
         ccode = ifelse(country == "Central African Rep.", 482, ccode),
         year = as.numeric(year)) %>%
  filter(year > 2014) %>%
  mutate(ccode1 = 365) %>% 
  dplyr::rename("exports_from_ccode1" = "exports", "ccode2" = "ccode") %>%
  dplyr::select("ccode1", "ccode2", "year", "exports_from_ccode1")


# IMF DOTS Imports from counterpart countries into US
imports.russia <- readxl::read_xlsx(here("../Data", "Exports_and_Imports_by_areas_and_Co (2).xlsx"), sheet = 1, skip = 5)

imports.russia <- imports.russia %>%
  dplyr::rename("country" = 1) %>%
  pivot_longer(cols = 2:ncol(imports.russia),
               values_to = "imports",
               names_to = "year") %>%
  filter(country %notin% netherlandsfilter) %>%
  mutate(ccode = countrycode(country, "country.name", "cown"),
         ccode = ifelse(country == "Eastern Germany", 265, ccode),
         ccode = ifelse(country == "China, P.R.", 710, ccode),
         ccode = ifelse(country == "Central African Rep.", 482, ccode),
         year = as.numeric(year)) %>%
  filter(year > 2014) %>% 
  mutate(ccode1 = 365) %>% 
  dplyr::rename("imports_to_ccode1" = "imports", "ccode2" = "ccode") %>%
  dplyr::select("ccode1", "ccode2", "year", "imports_to_ccode1")


# Combine US data
imf.russia <- exports.russia %>%
  full_join(imports.russia, by = c("ccode1", "ccode2", "year"))



#### Combine COW and IMF trade data for US and Russia ####
trade.data.final <- trade.com %>%
  bind_rows(imf.us) %>%
  bind_rows(imf.russia) %>%
  filter(!is.na(ccode2)) %>%
  arrange(ccode1, ccode2, year) %>% 
  mutate(tick = 1) %>% 
  group_by(ccode1, ccode2, year) %>% 
  dplyr::mutate(dups = cumsum(tick)) %>% 
  dplyr::arrange(-dups) %>% 
  filter(dups == 1) %>% 
  left_join(cpi, by = "year") %>%
  mutate(across(starts_with("imports|exports"), .x = .x/cpi))

# Check trade data comparison
tradecheck <- ggplot(data = trade.data.final %>% filter(ccode1 == 2 & ccode2 %in% c(20, 200, 220, 255, 356, 710, 732, 740, 900)), aes(x = year)) +
  geom_line(aes(y = exports_from_ccode1_cow), size = 2, color = "red") +
  geom_line(aes(y = exports_from_ccode1_g), size = 1.2, color = "blue") +
  geom_line(aes(y = exports_from_ccode1), size = 1.2, color = "black") +
  facet_wrap(. ~ ccode2)

tradecheck



# SIPRI Data
sipri.data <- readxl::read_xlsx(here::here("../Data/SIPRI-Milex-data-1949-2019.xlsx"), sheet = 7, skip = 4)

sipri.data <- sipri.data %>% 
  pivot_longer(cols = c(3:73),
               names_to = "year",
               values_to = "defburden") %>%
  filter(year >= 1950) %>% 
  mutate(ccode = countrycode(Country, "country.name", "cown")) %>% 
  mutate(ccode = ifelse(Country == "German DR", 265,
                        ifelse(Country == "Serbia", 345,
                               ifelse(Country == "Central African Rep.", 482, ccode)))) %>% 
  filter(defburden != "xxx" & defburden != ". .") %>% 
  mutate(defburden = round(as.numeric(defburden), digits = 4)) %>% 
  dplyr::select(ccode, year, defburden) %>% 
  mutate(year = as.numeric(year)) %>% 
  dplyr::rename("ccode2" = "ccode")

sipri.sp <- sipri.data 
  

# PRIO Data
# https://ucdp.uu.se/downloads/
prio.data <- read_csv(here("../Data/ucdp-prio-acd-191.csv")) %>% 
  rowwise() %>% 
  dplyr::select("conflict_id", "year", "intensity_level", "gwno_loc") %>% 
  separate(gwno_loc, sep = ",", into = c("ccode1", "ccode2", "ccode3", "ccode4" ,"ccode5", "ccode6")) %>% 
  pivot_longer(cols = c("ccode1", "ccode2", "ccode3", "ccode4", "ccode5", "ccode6"), 
               names_to = "countrynum", 
               values_to = "ccode", 
               values_drop_na = TRUE,
               names_repair = "minimal") %>% 
  mutate(ccode = as.numeric(trimws(ccode, which = "both")),
         intensity.low = if_else(intensity_level == 1, 1, 0),
         intensity.high = if_else(intensity_level == 2, 1, 0)) %>% 
  group_by(ccode, year) %>% # Gleditsch and Ward location country code and year
  dplyr::summarise(conflictcount = n_distinct(conflict_id),
            conflictcount_high = sum(intensity.high, na.rm = TRUE),
            conflictcount_low = sum(intensity.low, na.rm = TRUE),
            conflict_dummy = if_else(conflictcount > 0, 1, 0)) %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode2" = "ccode")


# PRIO data for spatial analysis
prio.sp <- prio.data 
  

# US War data using PRIO data
prio.wars <- read_csv(here("../Data/ucdp-prio-acd-191.csv")) %>% 
  rowwise() %>% 
  mutate(us_war = case_when(
    grepl(".*United States.*", side_a) ~ 1,
    grepl(".*United States.*", side_a_2nd) ~ 1,
    grepl(".*United States.*", side_b) ~ 1,
    grepl(".*United States.*", side_b_2nd) ~ 1,
    TRUE ~ 0
  ),
  russia_war = case_when(
    grepl(".*Russia|Soviet Union.*", side_a) ~ 1,
    grepl(".*Russia|Soviet Union.*", side_a_2nd) ~ 1,
    grepl(".*Russia|Soviet Union.*", side_b) ~ 1,
    grepl(".*Russia|Soviet Union.*", side_b_2nd) ~ 1,
    TRUE ~ 0
  )) %>% 
  filter(us_war == 1 | russia_war == 1) %>% 
  dplyr::select("location", "conflict_id", "year", "intensity_level", "gwno_loc", "us_war", "russia_war") %>% 
  separate(gwno_loc, sep = ",", into = c("ccode1", "ccode2", "ccode3", "ccode4" ,"ccode5", "ccode6")) %>% 
  pivot_longer(cols = c("ccode1", "ccode2", "ccode3", "ccode4", "ccode5", "ccode6"), 
               names_to = "countrynum", 
               values_to = "ccode", 
               values_drop_na = FALSE,
               names_repair = "minimal") %>% 
  mutate(ccode = as.numeric(trimws(ccode, which = "both"))) %>% 
  filter(!is.na(ccode) & ccode != 2) %>% 
  dplyr::select(year, ccode, us_war, russia_war) %>% 
  group_by(ccode, year) %>% 
  dplyr::summarise(us_war = max(us_war, na.rm = TRUE),
                   russia_war = max(russia_war, na.rm = TRUE)) %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode2" = "ccode")

prio.wars.sp <- prio.wars 


#### UN Ideal Point Data ####
undata <- read_delim(here("../Data", "Idealpoints.tab"), delim = "\t") %>% # Tab delimited
  dplyr::select(ccode, idealpoint, year) %>% 
  dplyr::filter(year >= 1950) %>% 
  mutate(ccode2 = list(seq(2:999))) %>% 
  unnest(ccode2)


undata.2 <- read_delim(here("../Data", "Idealpoints.tab"), delim = "\t") %>% # Tab delimited
  dplyr::select(ccode, idealpoint, year) %>% 
  dplyr::filter(year >= 1950) %>% 
  dplyr::rename("ccode2" = "ccode")


undata.com <- undata %>% 
  full_join(undata.2, by = c("ccode2", "year")) %>% 
  dplyr::arrange(ccode, ccode2, year) %>% 
  filter(!is.na(idealpoint.y)) %>% 
  filter(ccode2 == 2 | ccode2 == 365) %>% 
  mutate(idealdistance = abs(idealpoint.x - idealpoint.y)) %>% 
  dplyr::select(ccode, ccode2, year, idealdistance) %>% 
  pivot_wider(id_cols = c(ccode, year),
              values_from = "idealdistance",
              names_from = "ccode2",
              names_prefix = "idealdistance_") %>% 
  filter(ccode != 2) %>% 
  mutate(ccode = ifelse(ccode == 260, 255, ccode)) %>% 
  dplyr::rename("ccode2" = "ccode") 


#### ATOP Alliance Data ####
# ATOP alliance data
# Have to use a separate data set to isolate US and Russian allies
ally.data <- read_csv(here("../Data/ATOP V4.01 Data (csv)", "atop4_01ddyr.csv")) %>%
  dplyr::select(stateA, stateB, year, defense, offense) %>%
  filter(stateA %in% c(2, 365),
         year >= 1950) %>%
  mutate(us_ally = ifelse(defense == 1 & stateA == 2, 1, 0),
         russia_ally = ifelse(defense == 1 & stateA == 365, 1, 0),
         stateB = ifelse(stateB == 260, 255, stateB)) %>%
  dplyr::select(stateA, stateB, year, defense, us_ally, russia_ally) %>%
  dplyr::rename("ccode1" = "stateA", "ccode2" = "stateB") %>%
  group_by(ccode1, ccode2,  year) %>%
  dplyr::summarise(defense = max(defense, na.rm = TRUE),
                   us_ally = max(us_ally, na.rm = TRUE),
                   russia_ally = max(russia_ally, na.rm = TRUE)) # Get rid of Germany duplicate

ally.sp <- ally.data %>%
  dplyr::select(ccode1, ccode2, year, us_ally, russia_ally) %>%
  group_by(ccode2, year) %>%
  dplyr::summarise(us_ally = max(us_ally, na.rm = TRUE),
                   russia_ally = max(russia_ally, na.rm = TRUE))


#### Create mindist data ####

if (file.exists(here("../Data/mindist-dataframe.csv"))) {
  
  mindist <- read_csv(here("../Data/mindist-dataframe.csv"))
  
} else {

  tic()
  startdate = as.Date("1950-01-01", format = "%Y-%m-%d")
  enddate = as.Date("2016-01-01", format = "%Y-%m-%d")
  datelist <- data.frame(date = as.Date(seq(startdate, enddate, by = "1 year"), format = "%y-%m-%d"))

  thedate <- startdate

  mindist <- as.list(datelist$date) # Create empty list 

  mindist <- lapply(mindist, function(x) 
    df <- cshapes::distlist(date = x, type = "mindist", tolerance = 0.5, useGW = FALSE) %>% 
      mutate(year = x)
  )

  mindist <- bind_rows(mindist) %>% 
    mutate(year = format(year, "%Y"))

  write.csv(mindist, here("../Data/mindist-dataframe.csv"))

toc()

}

# Extend minimum distance to 2020
mindist.2 <- mindist %>% 
  filter(year >= 2013) %>% 
  mutate(year = year + 4)

mindist <- mindist %>% 
  bind_rows(mindist.2)


#### Distance to key locations #### 

strategic.locations <- c(365, 710, 731, 816, 812, 811, 645, 700, 345)

strategic.distance <- mindist %>% 
  arrange(ccode1, ccode2, year) %>% 
  filter(ccode2 %in% strategic.locations & ccode1 != 2) %>% 
  dplyr::rename("ccode" = "ccode1") %>% 
  group_by(ccode, year) %>% 
  pivot_wider(names_from = "ccode2",
              values_from = "mindist",
              names_prefix = "dist_") %>% 
  group_by(ccode, year) %>% 
  dplyr::summarise(across(.cols = starts_with("dist_"), ~max(.x, na.rm = TRUE))) %>% 
  dplyr::rename("ccode2" = "ccode")


#### Troop Data ####
# Need to correct Vietnam values. Troop data shows troops in Vietnam during the war, 
# but this should be South Vietnam.
troop.data <- get_troopdata(startyear = 1950, endyear = 2020, branch = TRUE) %>% 
  mutate(troops = ifelse(ccode == 816 & year <= 1975, 0, troops)) %>% 
  filter(ccode != 817)

troop.data.vietnam <- get_troopdata(host = 816, startyear = 1954, endyear = 1975, branch = TRUE) %>% 
  mutate(ccode = 817) %>% 
  dplyr::select(ccode, year, troops, army, navy, air_force, marine_corps) %>% 
  mutate(countryname = "Republic of Vietnam")

troop.data <- troop.data %>% 
  bind_rows(troop.data.vietnam) %>% 
  arrange(ccode, year) %>% 
  dplyr::rename("ccode2" = "ccode")
  
  
troop.data.sp <- troop.data 

#### Size of US Military over time #### 
# Create a time series frame with the total size of the US military by year
# Military size 1954-1993

milsize.1950.2005 <- readxl::read_xls(here::here("../Data", "troopMarch2005.xls"), sheet = 2) %>% # Kane's Heritage data.
  slice(6) %>% 
  dplyr::select(-1) %>% 
  mutate_if(is.numeric, as.character) %>% # Some conflicting types after reading data in
  group_by(Country) %>% 
  pivot_longer(cols = 2:57,
               values_to = "troops",
               names_to = "year") %>% 
  dplyr::select(Country, year, troops) %>% 
  mutate_each(funs = as.numeric) %>% # Convert all to numeric for append 
  filter(year <= 1993) 


# Military size 1994-2012

milsize <- readxl::excel_sheets(here::here("../Data", "AD_Strengths_FY1994-FY2012.xlsx"))

milsize.list <- lapply(milsize, function(x) readxl::read_excel(here::here("../Data", "AD_Strengths_FY1994-FY2012.xlsx"), sheet = x))

milsize.1.df <- bind_rows(milsize.list) %>% 
  filter(grepl(".*GRAND TOTAL.*", Rank)) %>% 
  dplyr::select("DoD Total") %>% 
  mutate(year = seq(1994, 2012)) %>% 
  dplyr::rename("Total" = "DoD Total") %>% 
  mutate(troops = as.numeric(Total),
         year = as.numeric(year)) %>% 
  dplyr::select(troops, year)

# Military size 2013-2016
milsize <- readxl::excel_sheets(here::here("../Data", "AD_Strengths_FY2013-FY2016.xlsx"))

milsize.list <- lapply(milsize, function(x) readxl::read_excel(here::here("../Data", "AD_Strengths_FY2013-FY2016.xlsx"), sheet = x))

milsize.2.df <- bind_rows(milsize.list) %>% 
  filter(grepl(".*GRAND TOTAL.*", `Department of Defense`)) %>% 
  dplyr::rename("Total" = 6) %>% 
  dplyr::select("Total") %>% 
  mutate(year = seq(2013, 2016)) %>% 
  mutate(troops = as.numeric(Total),
         year = as.numeric(year)) %>% 
  dplyr::select(troops, year)



# 2017 and 2018 data

# Troops 2017 (includes active duty only)
milsize.2017 <- readxl::read_xlsx(here::here("../Data", "DMDC_Website_Location_Report_1712.xlsx"), skip = 5) %>% 
  filter(grepl(".*GRAND.*", LOCATION)) %>% 
  dplyr::select(8) %>% 
  dplyr::rename("troops" = 1) %>% 
  mutate(year = 2017) %>% 
  mutate_each(as.numeric)


# Troops 2018 (includes active duty only)
milsize.2018 <- readxl::read_xlsx(here::here("../Data", "DMDC_Website_Location_Report_1812.xlsx"), skip = 5) %>% 
  filter(grepl(".*GRAND.*", LOCATION)) %>% 
  dplyr::select(8) %>% 
  dplyr::rename("troops" = 1) %>% 
  mutate(year = 2018) %>% 
  mutate_each(as.numeric)


milsize.1950.2018 <- bind_rows(milsize.1950.2005, milsize.1.df, milsize.2.df, milsize.2017, milsize.2018) %>% 
  dplyr::rename("usmilsize" = "troops") %>% 
  dplyr::select(year, usmilsize)



#### Major power deployments ####
troop.data.russia <- haven::read_dta(here("../Data/troopsComplete.dta")) %>% 
  filter(!(ccode == 255 & year == 1990)) %>% 
  mutate(ccode = ifelse(ccode == 260, 255, ccode)) %>% 
  arrange(ccode, year) %>% 
  dplyr::select(ccode, year, russiatroops) %>% 
  dplyr::rename("ccode2" = "ccode")



#### Inverse distance measures ####

# Conflict Weighted measure

spatial.measures <- mindist %>% 
  filter(ccode1 != ccode2) %>% 
  mutate(ccode1 = ifelse(ccode1 == 260, 255, ccode1),
         ccode2 = ifelse(ccode2 == 260, 255, ccode2),
         mindist = case_when(
           mindist < 1 | mindist == 0 ~ 1,
           TRUE ~ mindist)) %>%  # Note: distlist is returning weird values < 1 & > 0. These all appear to be contiguous.
  left_join(troop.data.sp, by = c("ccode2", "year")) %>%
  left_join(troop.data.russia, by = c("ccode2", "year")) %>% 
  left_join(prio.sp, by = c("ccode2", "year")) %>% 
  left_join(prio.wars.sp, by = c("ccode2", "year")) %>% 
  left_join(sipri.sp, by = c("ccode2", "year")) %>% 
  left_join(dem.sp, by = c("ccode2", "year")) %>% 
  left_join(gdp.sp, by = c("ccode2", "year")) %>% 
  left_join(ally.sp, by = c("ccode2", "year")) %>% 
  arrange(ccode1, ccode2, year) %>% 
  mutate(us_war = ifelse(is.na(us_war), 0, us_war),
         russia_war = ifelse(is.na(russia_war), 0, russia_war),
         conflictcount = ifelse(is.na(conflictcount), 0, conflictcount),
         conflictcount_high = ifelse(is.na(conflictcount_high), 0, conflictcount_high),
         conflictcount_low = ifelse(is.na(conflictcount_low), 0, conflictcount_low),
         conflict_dummy = ifelse(is.na(conflict_dummy), 0, conflict_dummy),
         inversedistance = 1 / mindist,
         troops_w = inversedistance * troops,
         russiatroops_w = inversedistance * russiatroops,
         us_war_w = inversedistance * us_war,
         russia_war_w = inversedistance * russia_war,
         us_ally_w = inversedistance * us_ally,
         russia_ally_w = inversedistance * russia_ally,
         conflictcount_w = inversedistance * conflictcount,
         conflictcount_high_w = inversedistance * conflictcount_high,
         conflictcount_low_w = inversedistance * conflictcount_low,
         conflict_dummy_w = inversedistance * conflict_dummy,
         v2x_polyarchy_w = inversedistance * v2x_polyarchy,
         polity2_w = inversedistance * polity2,
         defburden_w = inversedistance * defburden,
         gdp_w = inversedistance * wbgdp2011est) %>% 
  arrange(ccode1, ccode2, year) %>% 
  group_by(ccode1, year) %>% 
  dplyr::summarise(troops_w_sum = sum(troops_w, na.rm = TRUE),
                   troops_w_mean = mean(troops_w, na.rm = TRUE),
                   russiatroops_w_sum = sum(russiatroops_w, na.rm = TRUE),
                   russiatroops_w_mean = mean(russiatroops_w, na.rm = TRUE),
                   us_war_w_mean = mean(us_war_w, na.rm = TRUE),
                   russia_war_w_mean = mean(russia_war_w, na.rm = TRUE),
                   conflictcount_w_mean = mean(conflictcount_w, na.rm = TRUE),
                   conflictcount_high_w_mean = mean(conflictcount_high_w, na.rm = TRUE),
                   conflictcount_low_w_mean = mean(conflictcount_low_w, na.rm = TRUE),
                   conflict_dummy_w_mean = mean(conflict_dummy_w, na.rm = TRUE),
                   defburden_w_mean = mean(defburden_w, na.rm = TRUE),
                   v2x_polyarchy_w_mean = mean(v2x_polyarchy_w, na.rm = TRUE),
                   polity_w_mean = mean(polity2_w, na.rm = TRUE),
                   gdp_w_mean = mean(gdp_w, na.rm = TRUE),
                   us_ally_w_mean = mean(us_ally_w, na.rm = TRUE),
                   russia_ally_w_mean = mean(russia_ally_w, na.rm = TRUE)) %>% 
  mutate(troops_w_mean = ifelse(is.finite(troops_w_mean), troops_w_mean, NA),
         troops_w_sum = ifelse(is.finite(troops_w_sum), troops_w_sum, NA),
         us_war_w_mean = ifelse(is.finite(us_war_w_mean), us_war_w_mean, NA),
         conflictcount_w_mean = ifelse(is.finite(conflictcount_w_mean), conflictcount_w_mean, NA),
         conflictcount_high_w_mean = ifelse(is.finite(conflictcount_high_w_mean), conflictcount_high_w_mean, NA),
         conflictcount_low_w_mean = ifelse(is.finite(conflictcount_low_w_mean), conflictcount_low_w_mean, NA),
         conflict_dummy_w_mean = ifelse(is.finite(conflict_dummy_w_mean), conflict_dummy_w_mean, NA),
         defburden_w_mean = ifelse(is.finite(defburden_w_mean), defburden_w_mean, NA)) %>% 
  dplyr::rename("ccode2" = "ccode1")


data.clean <- dyads.complete %>% 
  left_join(gdp, by = c("ccode1", "year")) %>% 
  left_join(gdp.sp, by = c("ccode2", "year")) %>% 
  left_join(trade.data.final, by = c("ccode1", "ccode2", "year")) %>% 
  left_join(troop.data, by = c("ccode2", "year")) %>% 
  left_join(troop.data.russia, by = c("ccode2", "year")) %>% 
  left_join(undata.com, by = c("ccode2", "year")) %>% 
  left_join(ally.data, by = c("ccode1", "ccode2", "year")) %>% 
  left_join(prio.data, by = c("ccode2", "year")) %>% 
  left_join(prio.wars, by = c("ccode2", "year")) %>% 
  left_join(sipri.data, by = c("ccode2", "year")) %>% 
  left_join(milsize.1950.2018, by = c("year")) %>% 
  left_join(spatial.measures, by = c("ccode2", "year")) %>%
  left_join(strategic.distance, by = c("ccode2", "year")) %>% 
  dplyr::rename_at(vars(ends_with(".x")), ~glue::glue("{gsub('.x', '', .)}1")) %>% 
  dplyr::rename_at(vars(ends_with(".y")), ~glue::glue("{gsub('.y', '', .)}2")) %>% 
  group_by(ccode1, ccode2) %>% 
  mutate_at(vars("troops", "army", "air_force", "navy", "marine_corps", "russiatroops", "imports_to_ccode1", "exports_from_ccode1", "mindist", "wbpopest2", "troops_w_sum", "russiatroops_w_sum", "troops_w_mean", "russiatroops_w_mean"),
            list("lag" = ~dplyr::lag(.))) %>% 
  mutate_at(vars("troops", "army", "air_force", "navy", "marine_corps", "russiatroops", "imports_to_ccode1", "exports_from_ccode1", "mindist", "wbpopest2", contains("w_"), starts_with("dist_"), ends_with("lag")), 
            list("log" = ~log1p(.),
                 "max" = ~max(., na.rm = TRUE),
                 "cent" = ~(.) - mean(., na.rm = TRUE))) %>% 
  mutate_at(vars("troops_log"),
            list("z" = ~(((.) - mean(., na.rm = TRUE)) / (2*sd(., na.rm = TRUE))))) %>% 
  group_by(year) %>% 
  mutate_at(vars("imports_to_ccode1", "exports_from_ccode1", "wbgdp2011est2", "wbpopest2"),
            list("tcent" = ~(.) - mean(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  dplyr::mutate(time = year - 1950) %>% 
  group_by(ccode1, ccode2) %>%
  dplyr::mutate(c_time = year - min(year, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(atop_defense = factor(atop_defense, labels = c("Non-Ally", "Ally")),
         president = case_when(
           year <= 1952 ~ "Truman",
           year %in% c(1953:1960) ~ "Eisenhower",
           year %in% c(1961:1963) ~ "Kennedy",
           year %in% c(1964:1968) ~ "Johnson",
           year %in% c(1969:1973) ~ "Nixon",
           year %in% c(1974:1976) ~ "Ford",
           year %in% c(1977:1980) ~ "Carter",
           year %in% c(1981:1988) ~ "Reagan",
           year %in% c(1989:1992) ~ "Bush1",
           year %in% c(1993:2000) ~ "Clinton",
           year %in% c(2001:2008) ~ "Bush2",
           year %in% c(2009:2016) ~ "Obama",
           year %in% c(2017:2020) ~ "Trump",
           year %in% c(2021:2024) ~ "Biden"
         ),
         us_war = ifelse(is.na(us_war), 0, us_war)) %>% 
  dplyr::group_by(ccode1, ccode2, year) %>% 
  dplyr::mutate(tick = 1,
         dups = cumsum(tick)) 

write.csv(data.clean, file = here::here("../Data/data-predicting-troops.csv"))

  

```

